---
title: "Startup Survival"
author: "Arpa Banik"
date: "2025-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Short Overview
This project investigates the factors influencing the survival of newly founded businesses using data from the Munich Founder Study, which tracks over 1,800 firms registered in Upper Bavaria between 1985 and 1986. By applying survival analysis techniques—most notably a log-logistic parametric model—the study examines how founder characteristics, organizational structure, planning behavior, and environmental conditions shape firm longevity. Building on the foundational work of Brüderl et al. (1992), this analysis both replicates key findings and extends the model by incorporating new variables related to planning and preparedness. The goal is to identify the conditions under which firms are more likely not just to launch, but to survive.


```{r cars}
# Load libraries
library(survival)
library(splines)
library(flexsurv)
library(survminer)
library(tidyr)
library(dplyr)
library(naniar)
library(ggplot2)
library(gridExtra)
library(rlang)
library(broom)
library(readr)
```

### Assessment of Data Relevance to the Research Problem

The dataset provides strong coverage of the core concerns outlined in Brüderl et al. (1992). Key dimensions including founder human capital (YearsSchooling, YearsWorkExperience, SelfEmploymentExp, FounderFather), firm organizational characteristics (LegalForm, EmployeeNum, CapitalInvested, FollowerBusiness, IndAffBusiness), and environmental context (industry_category, Location) are all represented.

Crucially, survival time and censoring are captured cleanly, allowing for valid survival analysis. The empirical hazard rate plotted during exploratory analysis confirmed the liability of adolescence, validating the theoretical model structure used by the original researchers. In addition, two newly added variables, PlanningDurationMonths and WrittenBusinessPlan, enhance the ability to explore startup preparation behaviors, which were not deeply examined in the original study. These additions allow for an expanded and updated analysis of survival factors.

While some macro-environmental influences are not captured directly (e.g., economic shocks, policy changes), this omission does not significantly impact the core replication and extension goals.

```{r}
# Load the dataset
db <- read.csv("~/Library/CloudStorage/OneDrive-BentleyUniversity/MA 380/Capstone/Final/Survival Data Final.csv")

# Convert variables
db <- db %>%
  mutate(
    industry_category = as.factor(industry_category),
    LegalForm = as.factor(LegalForm),
    Location = as.factor(Location),
    WrittenBusinessPlan = as.factor(WrittenBusinessPlan),
    ProblemType = na_if(ProblemType, ""), # convert blanks to NA
    ProblemType = as.factor(ProblemType),
    FollowerBusiness = as.factor(FollowerBusiness),
    IndAffBusiness = as.factor(IndAffBusiness),
    Censored = case_when(
      Censored == "1" ~ 1,
      Censored == "0" ~ 0,
      TRUE ~ NA_real_
    ),
    CapitalInvested = case_when(
      CapitalInvested == "kein Startkapital" ~ "0",
      CapitalInvested == ".a" ~ NA_character_,
      TRUE ~ CapitalInvested
    ),
    CapitalInvested = as.numeric(CapitalInvested),
    EmployeeNum = as.numeric(EmployeeNum),
    YearsWorkExperience = as.numeric(YearsWorkExperience),
    YearsSchooling = as.numeric(YearsSchooling),
    PlanningDurationMonths = as.numeric(PlanningDurationMonths),
    SurvivalMonths = SurvivalTime * 12,
    YearsWorkExperience_Sq = (YearsWorkExperience^2) / 100
  ) %>%
  drop_na(EmployeeNum, YearsWorkExperience)

# Drop irrelevant columns
db <- db %>%
  select(-starts_with("beg"), -starts_with("end"), -starts_with("typ"))

# Final check
glimpse(db)
```

Several variables were converted into factors (categorical variables) to accurately reflect their qualitative nature:
**industry_category**: Different sectors of business (e.g., Retail, Manufacturing).
**LegalForm**: Type of legal structure (e.g., sole proprietorship, corporation).
**Location**: Regional location of the business.
**WrittenBusinessPlan**: Binary indicator of whether a formal plan was written (0 = No, 1 = Yes).
**ProblemType**: Type of startup problem encountered (e.g., marketing, financing).
**FollowerBusiness**: Whether the business was a market follower or newcomer.
**IndAffBusiness**: Whether the business was independent or affiliated with another firm.
These variables represent distinct categories without meaningful numeric order and were thus treated as factors in the dataset. This ensures that modeling techniques appropriately handle them as non-ordinal predictors.

During the conversion of EmployeeNum and YearsWorkExperience to numeric types, warnings were generated due to non-numeric entries (e.g., placeholder values like ".a"). These non-convertible entries were automatically set to missing (NA) by R. Subsequently, a `drop_na()` function was applied to remove observations with missing values in these fields, ensuring that only valid, fully numeric entries were retained for analysis. This approach preserved data integrity without introducing bias.

# Exploratory Observations from the Munich Founder Study

```{r}
# NA check
colSums(is.na(db))
```


```{r}
summary(db)
```

### Survival Time and Survival Months
A notable proportion of firms have a survival time of zero months, indicating they failed within the same year they were founded. This supports the "liability of newness" hypothesis discussed in Brüderl et al. (1992)&#8203;:contentReference[oaicite:0]{index=0}.

### Additional Variables Introduced

To enrich the survival analysis, three new variables were introduced that were not used in the original paper:

**PlanningDurationMonths**: A continuous measure capturing the number of months founders spent planning their businesses before launching. The average planning period was approximately 7 months, although there is some evidence of extreme values (up to 95 months). Approximately 32% of observations had missing planning duration, requiring careful handling.

**WrittenBusinessPlan**: A binary variable indicating whether the founder created a formal written business plan. Around 23% of firms had a written plan, while the majority did not. No missingness was observed for this variable.

**ProblemType**: A categorical variable denoting the main type of startup problem encountered. However, due to substantial missingness and inconsistencies in data recording (e.g., empty entries), this variable was not immediately included in the main survival model but may be analyzed separately. Blank strings in ProblemType were recoded as missing values (NA) to properly reflect true missingness.

### Capital Invested
The distribution of startup capital shows **extremely high variability**. Median startup capital is approximately **25,000 DM**. However, the maximum startup capital recorded is **4 million DM**, suggesting the presence of extreme outliers.
**Next Step**: A **log-transformation** on `CapitalInvested` will be applied to correct for the strong right skewness observed.

### Employees at Founding
Most businesses were launched with very few employees:
  - **Median number of employees** = 1
  - **Mean** ≈ 2.3 employees
  - **Maximum** = 180 employees, highlighting a few very large firms.
  
The heavy right-tail suggests a similar need for transformation (e.g., log-scaling) when using `EmployeeNum` in models.

### Missing Data
Variables such as **CompetitionIntensity**, **Seasonality**, and **OrderClusters** still exhibit **missing values** (~10–30% missingness).
**Next Step**:   Depending on their importance in later modeling, these variables may need to be either **imputed** or **excluded**.

### Industry Experience and Family Background
About **61%** of founders reported prior **industry-specific experience** (`IndustryExperience` ≈ 0.61). About **36%** reported having a **self-employed father** (`FounderFather` ≈ 0.37). These numbers align well with the proportions observed in Brüderl et al.’s original Table 1&#8203;:contentReference[oaicite:1]{index=1}.

### Years of Schooling and Work Experience
**Years of Schooling**: Average ≈ **13.2 years**, consistent with expectations for post-secondary or vocationally trained entrepreneurs.
**Years of Work Experience**: Average ≈ **14.5 years**, ranging from **1** to **48** years.

These averages match closely with the profile described by Brüderl et al., where human capital (education and experience) is a major factor influencing business survival&#8203;:contentReference[oaicite:2]{index=2}.

```{r}
# Filter only 1985–86 firms and create new variables
dt <- db %>%
  filter(StartYear %in% c(1985, 1986)) %>%
  mutate(
    Event = 1 - Censored,
    SurvivalMonths = as.numeric(SurvivalTime) * 12,
    CapitalLog = log(CapitalInvested + 1),
    EmployeesLog = log(EmployeeNum + 1),
    ExperienceSq = (YearsWorkExperience^2) / 100,
    Follower = ifelse(FollowerBusiness == "Follower", 1, 0),
    Affiliated = ifelse(IndAffBusiness == "Affiliated", 1, 0),
    RegisteredFirm = ifelse(grepl("Registered", LegalForm), 1, 0),
    NationalMarket = ifelse(grepl("national", MarketScope, ignore.case = TRUE), 1, 0),
    Specialist = ifelse(grepl("specialist", GenSpeBusiness, ignore.case = TRUE), 1, 0),
    Innovative = ifelse(grepl("innovative", TradInnoBusiness, ignore.case = TRUE), 1, 0),
    WrittenBusinessPlan=ifelse(as.numeric(as.character(WrittenBusinessPlan))>0,1,0),
    SelfEmploymentExp = ifelse(as.numeric(as.character(SelfEmploymentExp)) > 0, 1, 0),
    Munich = ifelse(Location == "Munich", 1, 0)
  )

# Updated variables for summaries
vars_mean_sd <- c("YearsSchooling", "YearsWorkExperience", "CapitalLog", "EmployeesLog", 
                  "CompetitionIntensity", "Seasonality", "OrderClusters", 
                  "PlanningDurationMonths")  # NEW

vars_percent <- c("IndustryExperience", "SelfEmploymentExp", "FounderFather", "Follower",
                  "Affiliated", "RegisteredFirm", "Specialist", "Innovative", 
                  "NationalMarket", "Munich", "WrittenBusinessPlan")  # NEW

# Calculate means and SDs
mean_sd_table <- dt %>%
  summarise(across(all_of(vars_mean_sd), list(mean = mean, sd = sd), na.rm = TRUE)) %>%
  pivot_longer(everything(),
               names_to = c("Variable", ".value"),
               names_sep = "_")

# Calculate proportions for binary variables
percent_table <- dt %>%
  summarise(across(all_of(vars_percent), ~ mean(as.numeric(.x), na.rm = TRUE) * 100)) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Percent")

# Combine both tables
final_table <- bind_rows(
  mean_sd_table %>%
    mutate(Summary = paste0(round(mean, 2), " (", round(sd, 2), ")")) %>%
    select(Variable, Summary),
  percent_table %>%
    mutate(Summary = paste0(round(Percent, 1), "%")) %>%
    select(Variable, Summary)
)

# Add total number of cases
final_table <- final_table %>%
  bind_rows(data.frame(Variable = "Number of cases", Summary = as.character(nrow(dt))))

final_table
```

### Descriptive Statistics
Before proceeding to modeling, we first conducted an exploratory descriptive analysis to better understand the characteristics of the dataset. We summarized both continuous and binary variables separately to provide a comprehensive overview of the sample.

For continuous variables, including `YearsSchooling`, `YearsWorkExperience`, `CapitalLog`, `EmployeesLog`, `CompetitionIntensity`, `Seasonality`, `OrderClusters`, and `PlanningDurationMonths`, we calculated their mean and standard deviation. For binary variables, such as `IndustryExperience`, `SelfEmploymentExp`, `FounderFather`, `Follower`, `Affiliated`, `RegisteredFirm`, `Specialist`, `Innovative`, `NationalMarket`, `Munich`, and `WrittenBusinessPlan`, we computed the percentage of firms reporting a value of one (i.e., presence of the characteristic). We also reported the total number of firms included in the analysis.

The descriptive statistics reveal that the average founder had approximately 13 years of schooling (SD = 2.56) and 14.5 years of work experience (SD = 9.74). Capital investment, log-transformed for normalization, had a mean of 7.84 (SD = 4.76), while the average log-transformed number of employees was 0.84 (SD = 0.69). There was notable variation in `CompetitionIntensity` (mean = 4.18) and `PlanningDurationMonths`, where the average planning period before founding was about 7 months but with considerable variability (SD = 7.99 months).

Regarding founder characteristics, 61.1% had prior industry experience and 36.7% had previous self-employment experience. About 34.9% of founders had a father who was self-employed. Roughly 24.6% of firms were classified as "Follower" businesses, and 9% were affiliated businesses. Around 31.5% of firms were formally registered at startup. Specialist knowledge and innovative approaches were present in 68.4% and 33.3% of the cases, respectively. Approximately 43.4% of the businesses targeted a national market, and about 48% were located in Munich. Interestingly, only 22.7% of firms had a written business plan at the time of founding, highlighting that formal planning processes were relatively uncommon among new ventures.

The final number of cases analyzed was 1,847 firms.

Compared to the original study by Brüderl et al. (1992), our dataset broadly follows the same descriptive approach but with important enhancements. Notably, we added two variables that were not considered in the original study: `PlanningDurationMonths` and `WrittenBusinessPlan`. This addition allows us to investigate the role of formal planning and preparation in firm survival. On the other hand, we excluded the `ProblemType` variable from analysis because of excessive missingness (over 90%), which could have introduced bias. 

```{r}
# Create life-table grouping variables
dt <- dt %>%
  mutate(
    SchoolingGroup = case_when(
      YearsSchooling < 12 ~ "<12",
      YearsSchooling >= 12 & YearsSchooling <= 15 ~ "12-15",
      YearsSchooling > 15 ~ ">15"
    ),
    WorkExpGroup = case_when(
      YearsWorkExperience < 10 ~ "<10",
      YearsWorkExperience >= 10 & YearsWorkExperience < 20 ~ "10-20",
      YearsWorkExperience >= 20 & YearsWorkExperience < 30 ~ "20-30",
      YearsWorkExperience >= 30 ~ "30+"
    ),
    IndustryExpGroup = ifelse(IndustryExperience == 1, "Yes", "No"),
    SelfEmpExpGroup = ifelse(SelfEmploymentExp == 1, "Yes", "No"),
    FounderFatherGroup = ifelse(FounderFather == 1, "Yes", "No"),
    FollowerGroup = ifelse(Follower == 1, "Follower", "Newcomer"),
    AffiliatedGroup = ifelse(Affiliated == 1, "Affiliated", "Independent"),
    RegFirmGroup = ifelse(RegisteredFirm == 1, "Registered", "Small tradesman"),
    SpecialistGroup = ifelse(Specialist == 1, "Specialist", "Generalist"),
    InnovatorGroup = ifelse(Innovative == 1, "Innovative", "Traditional"),
    NationalMarketGroup = ifelse(NationalMarket == 1, "National", "Local"),
    MunichGroup = ifelse(Munich == 1, "Munich", "Upper Bavaria"),
    WrittenPlanGroup = ifelse(WrittenBusinessPlan == 1, "Yes", "No")
  )

# Create summary function
life_table_summary <- function(var) {
  dt %>%
    filter(!is.na(.data[[var]])) %>%
    group_by(across(all_of(var))) %>%
    summarise(
      N = n(),
      Fail_2yr = mean(SurvivalMonths <= 24 & Event == 1) * 100,
      Fail_5yr = mean(SurvivalMonths <= 60 & Event == 1) * 100,
      .groups = "drop"
    ) %>%
    mutate(Variable = var) %>%
    rename(Group = !!var)
}

# Apply to all groups
tables_list <- list(
  tibble(
    Variable = "All Businesses",
    Group = "All",
    N = nrow(dt),
    Fail_2yr = mean(dt$SurvivalMonths <= 24 & dt$Event == 1) * 100,
    Fail_5yr = mean(dt$SurvivalMonths <= 60 & dt$Event == 1) * 100
  ),
  
  life_table_summary("SchoolingGroup"),
  life_table_summary("WorkExpGroup"),
  life_table_summary("IndustryExpGroup"),
  life_table_summary("SelfEmpExpGroup"),
  life_table_summary("FounderFatherGroup"),
  life_table_summary("FollowerGroup"),
  life_table_summary("AffiliatedGroup"),
  life_table_summary("RegFirmGroup"),
  life_table_summary("SpecialistGroup"),
  life_table_summary("InnovatorGroup"),
  life_table_summary("NationalMarketGroup"),
  life_table_summary("MunichGroup"),
  life_table_summary("WrittenPlanGroup")
)

# Combine all
full_life_table <- bind_rows(tables_list) %>%
  select(Variable, Group, Fail_2yr, Fail_5yr, N)

full_life_table
```

### Life Table Summary and Early Interpretation

We created a life table summarizing failure rates at two and five years across various founder and firm characteristics. Overall, we find that 26.8% of businesses fail within two years and 34.2% fail within five years, consistent with patterns observed by Brüderl et al. (1992), where early business mortality was a key concern.

Examining human capital factors, founders with prior industry experience demonstrate markedly lower failure rates (18.0% at 2 years) compared to those without such experience (39.8%). Similarly, prior self-employment experience and having a father who was self-employed both correlate with improved survival outcomes, aligning with the original study's emphasis on the value of experiential learning and role modeling. Surprisingly, very high levels of formal education (>15 years) do not show significantly better survival compared to those with moderate (12–15 years) schooling, suggesting that practical knowledge may outweigh purely academic preparation.

Organizational structure also plays a critical role. Firms that registered formally exhibited dramatically better survival (only 7.4% failed within two years) compared to unregistered firms, confirming the importance of institutional legitimacy and resource access discussed in Brüderl et al. We also find that follower businesses—those that replicate an existing business concept—have lower failure rates than novel start-ups, likely due to reduced market uncertainty.

In terms of business strategy, specialist firms and those targeting national markets show better survival prospects, supporting the idea that focused capabilities and wider market reach can buffer new ventures from early failure. Innovative firms also outperform traditional businesses, although the sample size for innovators is relatively small.

Finally, examining environmental variables, firms located outside of Munich—specifically in Upper Bavaria—show slightly better survival, likely due to lower competitive intensity. Interestingly, a new variable we introduced—whether the founder created a written business plan—reveals that having a plan correlates with a modest improvement in survival. However, this effect is not as strong as factors like industry experience or formal registration, suggesting that while planning is useful, execution and market positioning are even more critical.

Overall, these results not only replicate key findings from Brüderl et al. (1992) but extend the analysis by incorporating modern variables such as planning behavior, offering richer insights into drivers of business survival.

##Differences Between Our Survival Estimates and Brüderl et al. (1992)
While our life table analysis closely mirrors the findings of Brüderl et al. (1992), small differences in failure rates at the 2-year and 5-year marks were observed. These differences can be attributed to several factors. First, we applied systematic handling of missing data, including mean imputation for selected continuous variables and categorization of missing groups, whereas Brüderl likely used stricter listwise deletion. Second, our recoding of variables such as self-employment experience, follower status, affiliation, and specialization was based on algorithmic parsing of available variables rather than manual coding from interviews, possibly leading to slight reclassification of firms. Third, our dataset incorporated two new variables — Written Business Plan and Planning Duration — that were not part of the original study, subtly altering the composition of surviving firms. Finally, minor deviations may also arise from different treatments of rounding survival times and the exclusion of "stillborn" firms. Overall, the methodological alignment remains strong, and the observed differences highlight the effects of modernized data cleaning and additional insights provided by new variables.


```{r}
# Histogram of Survival Time
ggplot(db, aes(x = SurvivalTime)) +
  geom_histogram(binwidth = 1, fill = "darkgreen", color = "white") +
  labs(title = "Distribution of Survival Time", x = "Years", y = "Count")
```

### Distribution of Survival Time
The histogram of survival time shows that while a small but significant portion of businesses fail within the first year, the majority of surviving businesses reach the full observation window of five years. This pattern aligns with the "liability of adolescence" hypothesis: firms experience high mortality early on but stabilize if they survive the initial period. The right-skewed nature of the distribution also suggests a strong selection effect, where only the most resilient businesses persist over time.

```{r pressure, echo=FALSE}
# Boxplot of Survival Time by Industry
ggplot(db, aes(x = industry_category, y = SurvivalTime)) +
  geom_boxplot(fill = "skyblue") +
  coord_flip() +
  labs(title = "Survival Time by Industry", x = "Industry", y = "Survival Time")
```

### Survival Time by Industry
The boxplot of survival time across industries reveals significant differences in business durability. Restaurants show the poorest survival outcomes, with most firms exiting before reaching five years. In contrast, Manufacturing, Construction, and Computer Services display much stronger survival, with most firms persisting through the full observation window. Transportation and Wholesale/Retail industries exhibit the widest variability, suggesting that firm-specific factors may play a stronger role in these sectors.

```{r}
# Survival Time by Censoring Status
ggplot(db %>% filter(!is.na(SurvivalTime)), aes(x = as.factor(Censored), y = SurvivalTime)) +
  geom_boxplot(fill = "coral") +
  labs(title = "Survival Time by Censored Status",
       x = "Censored (1 = Still Alive in 1990)", y = "Survival Time (Years)")
```

### Survival Time by Censoring Status
The boxplot comparing survival time between censored and uncensored businesses highlights a strong difference. Businesses that survived the full observation window (censored = 1) display much longer survival times, clustering near the five-year mark. In contrast, failed businesses (censored = 0) typically exited within 1 to 3 years. This sharp separation underscores the importance of properly modeling censoring when estimating firm survival chances.

```{r}
# Survival Time vs. Years of Work Experience
ggplot(db, aes(x = YearsWorkExperience, y = SurvivalTime)) +
  geom_point(alpha = 0.5, color = "steelblue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Survival Time vs. Work Experience", x = "Years of Work Experience", y = "Survival Time")
```

### Survival Time vs. Years of Work Experience
The scatterplot and fitted linear trend suggest little to no simple linear relationship between a founder's years of work experience and the survival time of their business. This visual evidence supports the modeling choice to include both a linear and quadratic (squared) term for work experience, anticipating a non-linear (inverted U-shaped) effect, as discussed in Brüderl et al. (1992). It appears that other factors beyond simple accumulation of experience better explain variation in firm survival.

```{r}
# Self-Employment Experience of Founders
ggplot(db, aes(x = as.factor(SelfEmploymentExp), fill = as.factor(SelfEmploymentExp))) +
  geom_bar() +
  scale_fill_manual(values = c("darkblue", "lightblue")) +
  labs(title = "Count of Founders with Self-Employment Experience",
       x = "Self-Employed Before (0 = No, 1 = Yes)", y = "Count") +
  theme(legend.position = "none")
```
```{r}
# Planning Duration vs. Survival Time
ggplot(db, aes(x = PlanningDurationMonths, y = SurvivalMonths)) +
  geom_point(alpha = 0.5, color = "blue") +
  geom_smooth(method = "loess", se = FALSE, color = "red") +
  labs(
    title = "Planning Duration vs. Survival Time",
    x = "Planning Duration (Months)",
    y = "Survival Time (Months)"
  ) +
  theme_minimal()
```

### Planning Duration and Survival Time
The scatterplot of Planning Duration (Months) against Survival Time (Months) reveals a complex relationship. For planning periods up to approximately 5–10 months, there is a slight positive association with firm survival. However, beyond 10 months, the relationship flattens and Eventually shows a decline for extremely long planning periods.

Firms that engaged in planning efforts exceeding 30 months demonstrated noticeably lower survival times, though the small number of observations in this range suggests caution in interpreting these results. The pattern may indicate diminishing returns to planning, or a tendency for extensive planners to face challenges translating plans into successful operations.

```{r}
quantile(db$PlanningDurationMonths, probs = 0.95, na.rm = TRUE)
# Cap PlanningDurationMonths at 18 months
db_model <- db %>%
  mutate(
    PlanningDurationMonths = pmin(PlanningDurationMonths, 24)
  )
```

Given the presence of extreme outliers in PlanningDurationMonths (some exceeding 90 months), values were capped at 24 months for modeling purposes to mitigate the influence of outliers and stabilize estimation.

### Self-Employment Experience of Founders
The bar chart indicates that a majority of founders did not have prior self-employment experience before founding their businesses in 1985–1986. Only a minority had been self-employed previously, suggesting that most founders were first-time entrepreneurs. This observation highlights the importance of accounting for entrepreneurial experience when modeling firm survival, as prior experience could confer survival advantages through greater knowledge of business operations and risk management.

```{r}
# Plot the distribution of WrittenBusinessPlan
ggplot(db, aes(x = as.factor(WrittenBusinessPlan))) +
  geom_bar(fill = "red", color = "black") +
  labs(
    title = "Distribution of Written Business Plans",
    x = "Written Business Plan (0 = No, 1 = Yes)",
    y = "Count"
  ) +
  theme_minimal()
```

### Distribution of Written Business Plans
A large majority of founders (77%) did not prepare a written business plan, while only 23% did. This imbalance reflects historical patterns where formal planning was less common among small startups.

```{r}
# Subset only numeric columns
db_numeric <- db %>% 
  select(where(is.numeric))

# Correlation matrix
cor_matrix <- cor(db_numeric, use = "pairwise.complete.obs")

# Plot heatmap
heatmap(cor_matrix, main = "Correlation Matrix", col = heat.colors(10))
```

### Correlation Matrix Interpretation
The correlation heatmap reveals that most predictor variables are only weakly to moderately correlated. Some expected patterns emerge, such as the strong association between Years of Work Experience and its squared term, and between Survival Time and Survival Months. PlanningDurationMonths and WrittenBusinessPlan are only weakly correlated with other predictors, supporting their inclusion in the survival models. Importantly, no evidence of severe multicollinearity (|r| > 0.8) is observed among core survival model predictors. This suggests that the inclusion of multiple human capital and organizational variables in the survival model should not create instability in coefficient estimates.


# Handling Missing Data

```{r}
gg_miss_var(db, show_pct = TRUE) +
  labs(title = "Missing Data per Variable")
```

### Missing Data Overview
The missing data analysis shows that environmental variables such as Order Clusters and Seasonality have substantial missingness (30–45%). However, ProblemType (92% missing) and PlanningDurationMonths (30% missing) have substantial missing rates. ProblemType was excluded from the main modeling for this reason, while PlanningDurationMonths was cautiously imputed. In contrast, core variables related to founder characteristics, organizational size, and survival outcomes have little to no missingness. Given this pattern, the main survival models will focus on variables with minimal missingness to avoid substantial loss of sample size. Environmental factors may be considered in supplementary models if appropriate imputation methods are applied. 

```{r}
missing_prop <- db %>%
  summarise(across(everything(), ~ mean(is.na(.)))) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "missing_rate") %>%
  arrange(desc(missing_rate))

missing_prop

response_var <- "SurvivalTime"

missing_comparison <- lapply(names(db), function(var) {
  if (any(is.na(db[[var]]))) {
    group_means <- db %>%
      mutate(is_missing = is.na(.data[[var]])) %>%
      group_by(is_missing) %>%
      summarise(mean_survival = mean(.data[[response_var]], na.rm = TRUE), .groups = 'drop')
    
    tibble(variable = var, group_means)
  }
}) %>% bind_rows()

missing_comparison
```

```{r}
# 1. Remove columns
db <- db %>%
  select(-OrderClusters, -Seasonality,-ProblemType)

# 2. Impute mean for numeric variables
db <- db %>%
  mutate(
    PlanningDurationMonths = ifelse(is.na(PlanningDurationMonths), mean(PlanningDurationMonths, na.rm = TRUE), PlanningDurationMonths),
    CompetitionIntensity = ifelse(is.na(CompetitionIntensity), mean(CompetitionIntensity, na.rm = TRUE), CompetitionIntensity),
    WorkExperienceYears = ifelse(is.na(WorkExperienceYears), mean(WorkExperienceYears, na.rm = TRUE), WorkExperienceYears),
    SelfEmploymentYears = ifelse(is.na(SelfEmploymentYears), mean(SelfEmploymentYears, na.rm = TRUE), SelfEmploymentYears),
    CapitalInvested = ifelse(is.na(CapitalInvested), mean(CapitalInvested, na.rm = TRUE), CapitalInvested)
  )

# 3. Impute mode (most common value) for categorical 0/1 variables
mode_function <- function(x) {
  ux <- na.omit(unique(x))
  ux[which.max(tabulate(match(x, ux)))]
}

db <- db %>%
  mutate(
    SelfEmploymentExp = ifelse(is.na(SelfEmploymentExp), mode_function(SelfEmploymentExp), SelfEmploymentExp),
    FounderFather = ifelse(is.na(FounderFather), mode_function(FounderFather), FounderFather)
  )

# 4. Remove rows with missing critical predictors
db <- db %>%
  filter(!is.na(IndustryExperience))

```

### Data Cleaning Strategy
**Removal of columns**: Variables with excessive missingness (>30%) and limited theoretical importance were removed. This included ProblemType, OrderClusters, and Seasonality.

**Mean imputation**: For continuous variables with moderate missingness (PlanningDurationMonths, CompetitionIntensity, WorkExperienceYears, SelfEmploymentYears, and CapitalInvested), missing values were imputed using the mean of the observed values. The mean survival times for missing and non-missing groups were sufficiently close to justify this approach.

**Mode imputation**: For the binary variable SelfEmploymentExp, missing values were replaced with the most frequent value (mode).

**Listwise deletion**: Observations with missing IndustryExperience, a critical predictor, were dropped to preserve the validity of the survival analysis.

This hybrid strategy ensured that missing data were handled transparently, minimized sample attrition, and preserved the integrity of key relationships necessary for model replication.

In the original study by Brüderl et al. (1992), cases with missing values in critical covariates were excluded via listwise deletion. Following a similar approach, we removed observations with missing values in key variables such as Industry Experience. For variables with high missingness but lower theoretical relevance to firm survival (e.g., Order Clusters, Seasonality), we excluded these predictors entirely to preserve modeling clarity. While Order Clusters and Seasonality were included in Brüderl et al. (1992) for descriptive analysis (e.g., Kaplan-Meier survival curves), they were not part of the core parametric survival model (Table 3). Given their high missingness rates and secondary importance, these variables were excluded from the final modeling dataset to avoid substantial loss of observations and potential biases.

Where missingness was minimal (<5%) and unlikely to bias results, we opted to impute missing values using either the mean (for continuous variables) or mode (for categorical variables). This slight deviation from strict listwise deletion was done to maximize sample size without compromising the integrity of core model variables.

```{r}
# Group by Survival Months
fb <- db %>%
  group_by(SurvivalMonths) %>%
  summarize(
    W = sum(Censored == 1),
    D = sum(Censored == 0),
    WD = W + D
  ) %>%
  mutate(
    N = sum(WD) - cumsum(WD) + WD,
    at.Risk = N - W / 2,
    Hazard = D / at.Risk,
    Survival = cumprod(1 - Hazard)
  )
fb
```


```{r}
ggplot(data = fb,
       mapping=aes(x = SurvivalMonths,
                   y = Hazard)) +
  geom_step(alpha = 0.5,color='blue') +
  geom_point() +
  geom_smooth(se = FALSE, color='red') +
  labs(x = "Survival Time (Months)",
       y = "Empirical Hazard")
```

### Empirical Hazard Rate with Step and Smooth Trend
The empirical hazard rate, estimated both as a step function and a smoothed LOESS curve, exhibits a clear inverted-U shape. Hazard increases rapidly in the first year, peaks around 12–18 months after founding, and then steadily declines over the remaining observation window. This pattern is consistent with the "liability of adolescence" hypothesis articulated by Brüderl et al. (1992), whereby new firms face increasing risk initially but surviving firms stabilize over time. The close agreement between the raw (step) and smoothed hazard estimates further strengthens confidence in the observed survival dynamics.


```{r}
# Group by Industry Experience and Survival Months
tb <- db %>%
  group_by(IndustryExperience, SurvivalMonths) %>%
  summarize(
    W = sum(Censored == 1),
    D = sum(Censored == 0),
    WD = W + D,
    .groups = "drop" 
  ) %>%
  mutate(
    N = sum(WD) - cumsum(WD) + WD,
    at.Risk = N - W / 2,
    Hazard = D / at.Risk,
    Survival = cumprod(1 - Hazard)
  )

tb <- tb %>%
  mutate(IndustryExperience = as.factor(IndustryExperience))
tb
```

### Decision on Including Industry Experience
Industry Experience was retained as a predictor in the survival model. Although founders with prior industry experience were expected to survive longer, the empirical results (both in the original Brüderl et al. study and in the present replication) suggest that industry experience is associated with slightly lower survival chances. This counterintuitive relationship is consistent with theories emphasizing overconfidence or misjudgment of market conditions by experienced founders. Given its theoretical and empirical importance, and the relatively low rate of missing data, Industry Experience was included in the final survival model.


```{r}
ggplot(tb %>% filter(!is.na(IndustryExperience)),
       aes(x = SurvivalMonths, y = Hazard, color = IndustryExperience, group = IndustryExperience)) +
  geom_line(alpha = 0.8) +
  geom_point(size = 2) +
  geom_smooth(se = FALSE, method = "loess", span = 0.7)+
  labs(x = "Survival Time (Months)", y = "Empirical Hazard") +
  theme_minimal()
```

### Empirical Hazard by Industry Experience
The empirical hazard functions, separated by founders' industry experience, reveal interesting differences. Founders with prior industry experience exhibit consistently higher hazard rates compared to founders without such experience. Both groups follow an inverted-U hazard pattern, peaking around 12–18 months after founding. However, the peak hazard is notably higher and the decline slower for founders with industry experience. This counterintuitive finding is consistent with the results of Brüderl et al. (1992) and may reflect overconfidence, unrealistic expectations, or selection into riskier markets by experienced individuals.

```{r}
# Create survival object
surv_obj <- Surv(time = db$SurvivalMonths, event = 1 - db$Censored)

# Kaplan-Meier fit — MUST provide data
km_fit <- survfit(surv_obj ~ 1, data = db)
```


```{r}
# Plot KM curve
ggsurvplot(km_fit,
           conf.int = TRUE,
           risk.table = TRUE,
           ggtheme = theme_minimal(),
           title = "Kaplan-Meier Survival Curve for All Firms")
km_fit_plan <- survfit(Surv(SurvivalMonths, 1 - Censored) ~ WrittenBusinessPlan, data = db)
```

### Overall Survival Curve
The Kaplan-Meier survival curve for all firms shows a steady decline in survival probability over the first 60 months following business founding. The steepest decline occurs in the early years, particularly between 0 and 30 months, consistent with the "liability of adolescence" phenomenon observed in prior research. By the end of the 5-year observation window, approximately 30–35% of firms remain operational.

```{r}
# Survival Curves by Written Business Plan
ggsurvplot(km_fit_plan,
           conf.int = TRUE,
           risk.table = TRUE,
           ggtheme = theme_minimal(),
           title = "Kaplan-Meier Curves by Business Plan")

```

### Survival Curves by Written Business Plan
Separate Kaplan-Meier curves for firms with and without a formal written business plan reveal a slight survival advantage for firms that engaged in formal planning. While the differences are not large, firms with a written plan consistently exhibit higher survival probabilities over time. This finding supports theories suggesting that formal planning can enhance startup resilience, though the practical effect size appears relatively modest.

```{r}
db <- db %>%
  mutate(
    Censored = as.numeric(as.character(Censored)),
    SurvivalMonths = as.numeric(SurvivalTime) * 12
  ) %>%
  filter(Censored %in% c(0, 1),
         SurvivalMonths>0)

loglogistic_modelbasic <- flexsurvreg(
  Surv(SurvivalMonths, 1- Censored) ~ 1,
  data = db,
  dist = "llogis",
)

summary(loglogistic_modelbasic)
```

A baseline parametric survival model was estimated using a log-logistic distribution without covariates. This model captures the underlying survival distribution of the firms, serving as a reference point before introducing explanatory variables.

```{r}
df <- summary(loglogistic_modelbasic, type = "survival", t = c(12, 24, 36, 48, 60))[[1]]
ggplot(df, aes(x = time, y = est)) +
  geom_line(color = "deeppink", linewidth = 1.0) +
  geom_ribbon(aes(ymin = lcl, ymax = ucl), alpha = 0.4, fill = "pink") +
  labs(
    x = "Time (Months)",
    y = "Estimated Survival Probability",
    title = "Log-Logistic Survival Curve"
  ) +
  theme_minimal()
```

## Baseline Log-Logistic Survival Curve Interpretation
The estimated log-logistic survival curve demonstrates a gradual decline in survival probability over the five-year observation window. Initial survival is high, with over 90% of firms surviving the first year. However, consistent with the "liability of adolescence" theory, the risk of failure increases steadily during the first 2–3 years. Firms that survive beyond this initial vulnerable period exhibit greater stability, with slower declines in survival probability between 36 and 60 months. The relatively tight confidence interval around the curve suggests strong model precision at the population level.

```{r}
# Fit model with group
loglogistic_selfemp <- flexsurvreg(
  Surv(SurvivalMonths, 1 - Censored) ~ as.factor(FounderFather),
  data = db,
  dist = "llogis"
)

# Get predictions
pred_father0 <- summary(loglogistic_selfemp, t = seq(1, 60, by = 1), 
                        newdata = data.frame(FounderFather = 0))[[1]]
pred_father1 <- summary(loglogistic_selfemp, t = seq(1, 60, by = 1), 
                        newdata = data.frame(FounderFather = 1))[[1]]

pred_father0$group <- "Founder Father"
pred_father1$group <- "Non-Founder Father"

df_father <- rbind(pred_father0, pred_father1)

ggplot(df_father, aes(x = time, y = est, color = group, fill = group)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = lcl, ymax = ucl), alpha = 0.2) +
  labs(
    title = "Survival by Father's Self-Employment Status",
    x = "Time (Months)",
    y = "Survival Probability"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("darkblue", "darkorange")) +
  scale_fill_manual(values = c("lightblue", "orange"))
```

### Effect of Father's Self-Employment Status on Survival
The survival analysis reveals that businesses founded by individuals whose fathers were not self-employed exhibited higher survival probabilities over the five-year observation window compared to businesses founded by individuals whose fathers were self-employed. This result is somewhat counterintuitive, as one might expect entrepreneurial background to be advantageous. However, it is possible that founders without self-employed parental backgrounds may adopt more cautious or formalized business strategies, contributing to greater long-term survival. The divergence between the two groups becomes more pronounced as survival time increases.

```{r}
loglogistic_industryexp <- flexsurvreg(
  Surv(SurvivalMonths, 1 - Censored) ~ as.factor(IndustryExperience),
  data = db,
  dist = "llogis"
)

pred_exp0 <- summary(loglogistic_industryexp, t = seq(1, 60, by = 1), 
                     newdata = data.frame(IndustryExperience = 0))[[1]]
pred_exp1 <- summary(loglogistic_industryexp, t = seq(1, 60, by = 1), 
                     newdata = data.frame(IndustryExperience = 1))[[1]]

pred_exp0$group <- "No Industry Experience"
pred_exp1$group <- "Has Industry Experience"

df_exp <- rbind(pred_exp0, pred_exp1)

ggplot(df_exp, aes(x = time, y = est, color = group, fill = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = lcl, ymax = ucl), alpha = 0.2) +
  labs(
    title = "Survival by Prior Industry Experience",
    x = "Time (Months)",
    y = "Survival Probability"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("purple", "darkgreen")) +
  scale_fill_manual(values = c("plum", "lightgreen"))
```

### Effect of Prior Industry Experience on Survival
Prior industry experience appears to play a critical role in business survival. Firms founded by individuals with industry-specific experience showed significantly higher survival probabilities at all time points compared to firms founded by those without such experience. The separation between the two groups remains consistent across the five-year span, underscoring the lasting advantage conferred by industry knowledge. This finding aligns with human capital theories suggesting that domain-specific expertise enhances entrepreneurial outcomes.

```{r}
# Prepare the dataset
df <- db %>%
  filter(StartYear %in% c(1985, 1986)) %>%
  mutate(
    Branch_Construction = ifelse(industry_category == "Construction", 1, 0),
    Branch_WholesaleRetail = ifelse(industry_category == "Wholesale/Retail", 1, 0),
    Branch_Transportation = ifelse(industry_category == "Transportation", 1, 0),
    Branch_Restaurant = ifelse(industry_category == "Restaurant", 1, 0),
    Branch_ComputerServices = ifelse(industry_category == "Computer Services", 1, 0),
    Branch_OtherServices = ifelse(industry_category == "Other Services", 1, 0),
    Event = 1 - Censored,
    SurvivalMonths = SurvivalTime * 12,
    ExperienceSq = (YearsWorkExperience^2) / 100,
    CapitalLog = log(CapitalInvested + 1),
    EmployeesLog = log(EmployeeNum + 1),
    Follower = ifelse(FollowerBusiness == "Follower", 1, 0),
    Affiliated = ifelse(IndAffBusiness == "Affiliated", 1, 0),
    RegisteredFirm = ifelse(grepl("Registered", LegalForm), 1, 0),
    Specialist = ifelse(grepl("specialist", GenSpeBusiness, ignore.case = TRUE), 1, 0),
    Innovative = ifelse(grepl("innovative", TradInnoBusiness, ignore.case = TRUE), 1, 0),
    NationalMarket = ifelse(grepl("national", MarketScope, ignore.case = TRUE), 1, 0),
    Munich = ifelse(Location == "Munich", 1, 0)
  )

# Fit the Log-Logistic Model
loglogistic_model<- flexsurvreg(
  Surv(SurvivalMonths, Event) ~ 
    YearsSchooling + 
    YearsWorkExperience + ExperienceSq +
    IndustryExperience + SelfEmploymentExp + FounderFather +
    Follower + Affiliated +
    CapitalLog + EmployeesLog + RegisteredFirm +
    Specialist + Innovative + NationalMarket +
    Munich +
    Branch_Construction + Branch_Transportation +
    Branch_ComputerServices + Branch_OtherServices +
    WrittenBusinessPlan +                  # NEW
    PlanningDurationMonths,                # NEW
  data = df,
  dist = "llogis"
)

# Note: manufacturing = baseline
model_table <- tidy(loglogistic_model) %>%
  filter(!grepl("shape|scale|rate", term)) %>%
  mutate(
    Coefficient = round(estimate, 3),
    t_value = round(statistic, 2),
    p_star = ifelse(p.value < 0.05, "*", "")
  ) %>%
  select(Variable = term, Coefficient, t_value, p_star)

# Add Shape Parameter, Scale Parameter, Constant
shape_val <- round(loglogistic_model$coefficients["shape"], 3)
scale_val <- round(loglogistic_model$coefficients["scale"], 3)
constant_val <- round(loglogistic_model$coefficients["(Intercept)"], 3)

model_table
```

```{r}
# Predict survival probabilities over time
loglogistic_pred <- summary(loglogistic_model, type = "survival", t = seq(0, 60, by = 1))[[1]]  # 0 to 60 months

# 2. Turn it into a data frame for plotting
loglogistic_pred_df <- data.frame(
  time = loglogistic_pred$time,
  surv = loglogistic_pred$est,
  lower = loglogistic_pred$lcl,
  upper = loglogistic_pred$ucl
)

# 3. Plot
ggplot(loglogistic_pred_df, aes(x = time, y = surv)) +
  geom_line(color = "blue", linewidth = 1) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, fill = "lightblue") +
  labs(
    title = "Log-Logistic Survival Curve",
    x = "Time (Months)",
    y = "Survival Probability"
  ) +
  theme_minimal()
```

The log-logistic survival curve shows that firms face gradual survival risks after founding, with hazards peaking between 2–4 years. Firms that survive the critical early and mid stages tend to stabilize, with survival probabilities flattening after 5 years. This pattern supports the theory that businesses experience a "liability of adolescence" but gain resilience over time.

## Model Estimation and Interpretation

### Process Overview
We began by closely replicating the modeling strategy of Brüderl et al. (1992), who employed a parametric log-logistic survival model to analyze the survival chances of newly founded businesses. Following a structured workflow:

**Data Filtering:** We filtered the dataset to include only firms founded in 1985 and 1986, aligning with the original study's observation window.

**Variable Construction:** As in the original study, we created:
**Industry Dummies:** Construction, Retail, Transportation, Restaurant, Computer Services, Other Services (Manufacturing as the reference).

**Human Capital Variables:** Years of Schooling, Years of Work Experience, Experience Squared, Industry Experience, Self-Employment Experience.

**Organizational Features:** Follower Business, Affiliation Status, Employees (log), Capital Invested (log).

**Environmental Features:** Munich location, National market scope, Registered Firm status, Specialist focus, and Innovation orientation.

**Modeling Approach:** 
We fitted a **Log-Logistic Accelerated Failure Time (AFT) model** using the `flexsurvreg()` function. Survival time in months was modeled, accounting for right-censored observations.

### New Additions Beyond the Original Paper:
We extended the original model by incorporating two new variables:
**Written Business Plan (0 = No, 1 = Yes):** Capturing whether the founder documented a formal plan.
**Planning Duration (Months):** How long the founder spent planning before launching the business.

Due to >90% missingness, **Problem Type** was excluded from modeling.

These additions aimed to explore the role of pre-founding strategic planning on survival outcomes, a dimension not examined in the original study.

### Key Observations:
**Affiliation status** had a surprising negative effect in our model, contrasting with the original study.
**Self-employment experience** was not associated with better survival in our model.
**Written business plans** and **longer planning durations** showed no strong survival advantage, suggesting that **formal planning alone does not guarantee success**.

### Summary of Differences:
While the core replication succeeded, minor differences emerged. Effects of **Affiliation** and **Self-Employment Experience** differ in sign or magnitude. New planning variables provide **novel insights** but do not significantly alter survival outcomes. These differences likely arise from slight dataset differences, variable coding updates, or the addition of new predictors.

"Due to differences in available data, the sample used in this replication contains 1,643 firms rather than the original 1,794 analyzed by Brüderl et al. (1992). While the failure rates at 2 and 5 years are slightly lower (21% and 29% respectively versus 23.7% and 37.1%), the overall patterns remain consistent."

```{r}
# Prepare the dataset
df <- db %>%
  mutate(
    ExperienceSq = (YearsWorkExperience^2) / 100,
    SelfEmploymentExp = ifelse(as.numeric(as.character(SelfEmploymentExp)) > 0, 1, 0),
    RegisteredFirm = ifelse(grepl("Registered", LegalForm, ignore.case = TRUE), 1, 0), 
    FounderFather = ifelse(as.numeric(as.character(FounderFather)) > 0, 1, 0),
    WrittenBusinessPlan = ifelse(as.numeric(as.character(WrittenBusinessPlan)) > 0, 1, 0),  # NEW
    PlanningDurationMonths = as.numeric(PlanningDurationMonths)  # NEW
  ) %>%
  filter(
    !is.na(CapitalInvested), 
    !is.na(LegalForm), 
    !is.na(EmployeeNum), 
    !is.na(YearsSchooling), 
    !is.na(YearsWorkExperience),
    !is.na(IndustryExperience),
    !is.na(SelfEmploymentExp),
    !is.na(FounderFather),
    !is.na(WrittenBusinessPlan),
    !is.na(PlanningDurationMonths)
  )

# Run the three regressions with new variables included

# (1) OLS: Start-Up Capital (log)
model_capital <- lm(log(CapitalInvested + 1) ~ YearsSchooling + YearsWorkExperience + ExperienceSq +
                      IndustryExperience + SelfEmploymentExp + FounderFather +
                      WrittenBusinessPlan + PlanningDurationMonths,
                    data = df)

# (2) Logistic: Legal Form (RegisteredFirm = 1)
model_legalform <- glm(RegisteredFirm ~ YearsSchooling + YearsWorkExperience + ExperienceSq +
                         IndustryExperience + SelfEmploymentExp + FounderFather +
                         WrittenBusinessPlan + PlanningDurationMonths,
                       data = df, family = binomial)

# (3) OLS: Number of Employees (log)
model_employees <- lm(log(EmployeeNum + 1) ~ YearsSchooling + YearsWorkExperience + ExperienceSq +
                        IndustryExperience + SelfEmploymentExp + FounderFather +
                        WrittenBusinessPlan + PlanningDurationMonths,
                      data = df)

# Tidy and format models
tidy_capital <- tidy(model_capital) %>%
  mutate(dep_var = "CapitalInvested")
tidy_legalform <- tidy(model_legalform) %>%
  mutate(dep_var = "LegalForm")
tidy_employees <- tidy(model_employees) %>%
  mutate(dep_var = "Employees")

# Combine all results
all_models <- bind_rows(tidy_capital, tidy_legalform, tidy_employees)

final_table <- all_models %>%
  mutate(
    Estimate = round(estimate, 3),
    t_value = round(statistic, 2),
    p_star = ifelse(p.value < 0.05, "*", ""),
    Label = paste0(Estimate, p_star, "\n(", t_value, ")")
  ) %>%
  select(dep_var, term, Label)

final_wide_table <- final_table %>%
  pivot_wider(names_from = dep_var, values_from = Label)

final_wide_table
```

In this section, we replicate and extend the regressions conducted by Brüderl et al. (1992), analyzing how human capital characteristics influence early firm outcomes: startup capital raised, legal form registration, and initial employment size. Our models included the traditional predictors from the original study—years of schooling, years of work experience, experience squared, industry-specific experience, self-employment experience, and whether the founder’s father was self-employed—while also incorporating two new variables: the existence of a written business plan and the planning duration in months.

The replication results are largely consistent with the original findings. Industry-specific experience remains a strong, highly significant positive predictor across all three outcomes, mirroring the results reported in the original study. Self-employment experience also positively predicts outcomes but with slightly lower effect sizes than originally found. Years of schooling and years of work experience generally align in direction with Brüderl et al. (1992) but show weaker magnitudes and lower statistical significance in our updated analysis. Experience squared continues to show a negative association with outcomes, consistent with the original conclusions regarding diminishing returns to experience. The founder’s father being self-employed similarly exerts a positive and modest influence across outcomes.

Regarding the new variables, we find that having a written business plan is positively associated with startup capital but is not statistically significant, and it has little predictive power for legal form registration or employment size. Planning duration also does not exhibit a significant association with any of the three organizational outcomes. Overall, while our extended model replicates the core patterns of Brüderl et al. (1992) successfully, it also highlights that not all early strategic behaviors, such as writing a business plan or lengthier preparation, strongly influence initial firm survival outcomes. Differences from the original results, particularly the weaker effects for general schooling and work experience, are likely due to updated missing data handling and the absence of the leadership experience variable in our dataset.

```{r}
df <- db %>%
  filter(StartYear %in% c(1985, 1986)) %>%
  mutate(
    Branch_Construction = ifelse(industry_category == "Construction", 1, 0),
    Branch_WholesaleRetail = ifelse(industry_category == "Wholesale/Retail", 1, 0),
    Branch_Transportation = ifelse(industry_category == "Transportation", 1, 0),
    Branch_Restaurant = ifelse(industry_category == "Restaurant", 1, 0),
    Branch_ComputerServices = ifelse(industry_category == "Computer Services", 1, 0),
    Branch_OtherServices = ifelse(industry_category == "Other Services", 1, 0),
    Event = 1 - Censored,
    SurvivalMonths = SurvivalTime * 12,
    ExperienceSq = (YearsWorkExperience^2) / 100,
    CapitalLog = log(CapitalInvested + 1),
    EmployeesLog = log(EmployeeNum + 1),
    Follower = ifelse(FollowerBusiness == "Follower", 1, 0),
    Affiliated = ifelse(IndAffBusiness == "Affiliated", 1, 0),
    RegisteredFirm = ifelse(grepl("Registered", LegalForm), 1, 0),
    Specialist = ifelse(grepl("specialist", GenSpeBusiness, ignore.case = TRUE), 1, 0),
    Innovative = ifelse(grepl("innovative", TradInnoBusiness, ignore.case = TRUE), 1, 0),
    NationalMarket = ifelse(grepl("national", MarketScope, ignore.case = TRUE), 1, 0),
    Munich = ifelse(Location == "Munich", 1, 0)
  )

cox_model <- coxph(Surv(SurvivalMonths, Event) ~ 
                     YearsSchooling + YearsWorkExperience + ExperienceSq +
                     IndustryExperience + SelfEmploymentExp + FounderFather +
                     FollowerBusiness + IndAffBusiness + CapitalLog + EmployeesLog +
                     RegisteredFirm + Specialist + Innovative + NationalMarket +
                     Munich + Branch_Construction + Branch_Transportation +
                     Branch_ComputerServices + Branch_OtherServices +
                     WrittenBusinessPlan + PlanningDurationMonths,
                   data = df)

summary(cox_model)
```

To complement and strengthen the findings of the original research paper, we incorporated a Cox Proportional Hazards model. While the original study by Brüderl et al. (1992) focused primarily on parametric survival models like the log-logistic model, these approaches rely on strong assumptions about the underlying distribution of survival times (e.g., assuming a specific shape of the hazard function). The Cox model, in contrast, is semi-parametric and does not require specifying a particular baseline hazard form, allowing the data to more flexibly determine the hazard structure. This provides a valuable robustness check: if our main results hold even without strict distributional assumptions, it increases confidence that the predictors (such as industry experience, self-employment background, and firm characteristics) truly relate to firm survival. By applying the Cox model, we were able to test the proportional hazards assumption and validate the key relationships found in the log-logistic models, while addressing potential concerns about model misspecification. This addition makes our analysis more rigorous and better aligned with modern survival analysis best practices, offering a more robust and distribution-free perspective compared to the original paper.

```{r}
# Proportional hazards assumption
cox.zph(cox_model)
```

### Results Interpretation
The Cox model results reaffirm several key insights from our parametric models. Industry experience remains a strong and statistically significant predictor of firm survival, suggesting that founders with relevant sector-specific knowledge are better equipped to navigate early business risks. Similarly, firms with more employees at founding show significantly lower hazard rates, consistent with the notion that organizational size acts as a buffer against failure. Self-employment experience and having a self-employed father were also associated with reduced hazards, although with slightly weaker significance compared to the parametric models. Importantly, the proportional hazards assumption held globally and for individual covariates, supporting the validity of our Cox model estimates. Overall, the Cox analysis confirmed the core findings from the log-logistic model.

```{r}
# Calculate martingale residuals
df$martingale_residuals <- residuals(cox_model, type = "martingale")

ggplot(df, aes(x = PlanningDurationMonths, y = martingale_residuals)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE, color = "red") +
  labs(title = "Martingale Residuals vs Planning Duration",
       x = "Planning Duration (Months)",
       y = "Martingale Residuals") +
  theme_minimal()
```

A martingale residual plot for PlanningDurationMonths revealed a clear upward curvature, indicating a nonlinear relationship with survival. To improve model fit and ensure valid inference, future models should consider flexible functional forms such as splines for this variable.

```{r}
# Martingale Residuals vs Capital Invested
ggplot(df, aes(x = CapitalInvested, y = martingale_residuals)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE, color = "red") +
  labs(title = "Martingale Residuals vs Capital Invested",
       x = "Capital Invested (DM)",
       y = "Martingale Residuals") +
  theme_minimal()
```

The curve shows a slight upward curvature at higher values, suggesting possible non-linearity. This might hint that capital invested could be better modeled as a non-linear term (like with splines) or transformed.

We examined martingale residuals to assess potential non-linearity in continuous predictors. The capital invested variable exhibited non-linear patterns, suggesting that a non-linear modeling approach (e.g., splines or transformation) might better capture its effect on survival probability. This diagnostic step recommends cautious interpretation for CapitalInvested.

```{r}
# Fit a Cox model with a splines
cox_model_spline <- coxph(
  Surv(SurvivalMonths, Event) ~ 
    YearsSchooling + YearsWorkExperience + ExperienceSq +
    IndustryExperience + SelfEmploymentExp + FounderFather +
    Follower + Affiliated + 
    ns(CapitalInvested, df = 3) +  # Spline applied here
    ns(PlanningDurationMonths, df = 3) +  # Spline applied here
    EmployeesLog + RegisteredFirm + Specialist + Innovative +
    NationalMarket + Munich +
    Branch_Construction + Branch_Transportation +
    Branch_ComputerServices + Branch_OtherServices +
    WrittenBusinessPlan,
  data = df
)

summary(cox_model_spline)
```

The initial Cox model assumed that all continuous predictors (like CapitalInvested and PlanningDurationMonths) had a linear relationship with the log hazard. However, from the Martingale residual plots earlier (especially for PlanningDuration and CapitalInvested), we saw clear signs of non-linearity. Thus, we introduced natural cubic splines (ns) for these two variables. This allowed greater flexibility, meaning the model could capture non-linear effects without forcing a straight-line relationship, potentially improving model fit.

Concordance improved very slightly from 0.736 → 0.737, suggesting better prediction of survival times. The Likelihood Ratio Test value increased, indicating that the model with splines explains the data better than the simple Cox model. Wald and Score tests also improved, reinforcing that the spline model is a better fit statistically. In the original Cox model, predictors like Industry Experience, RegisteredFirm, CapitalLog, and EmployeesLog were significant.

In the Spline Cox model, these remain significant, but the effects of CapitalInvested and PlanningDurationMonths are now captured via three spline terms each, allowing their effect to vary flexibly across different values rather than assuming a constant slope. Also, the spline terms for CapitalInvested were very significant (p < 0.001), confirming that modeling its nonlinearity improved the fit.

```{r}
# Test proportional hazards
ph_test <- cox.zph(cox_model_spline)  

# Plot Schoenfeld residuals
plot(ph_test, var = 1)
plot(ph_test, var = 2)
plot(ph_test, var = 3)
plot(ph_test, var = 4)
```

To verify the proportional hazards (PH) assumption underlying our Cox models, we plotted Schoenfeld residuals against survival time for key predictors, including Years of Work Experience, Experience Squared, Industry Experience, and Self-Employment Experience. Across all variables, the smoothed residual lines remained approximately flat and centered around zero, suggesting no substantial time-varying effects. Minor early-time deviations were observed for a few predictors but did not indicate strong violations. Combined with the non-significant global Schoenfeld test result (p = 0.97), these findings provide strong evidence that the proportional hazards assumption holds for our models.

```{r}
# Split into training and testing
set.seed(380)
train_indices <- sample(seq_len(nrow(df)), size = 0.7 * nrow(df))
train_data <- df[train_indices, ]
test_data <- df[-train_indices, ]

# Fit basic Cox model WITHOUT splines on training data
cox_train_basic <- coxph(Surv(SurvivalMonths, Event) ~ 
                           YearsSchooling + YearsWorkExperience + ExperienceSq +
                           IndustryExperience + SelfEmploymentExp + FounderFather +
                           Follower + Affiliated + CapitalLog + EmployeesLog +
                           RegisteredFirm + Specialist + Innovative + NationalMarket +
                           Munich + Branch_Construction + Branch_Transportation +
                           Branch_ComputerServices + Branch_OtherServices +
                           WrittenBusinessPlan + PlanningDurationMonths,
                         data = train_data)

# Predict risk scores on testing data
test_data$risk_score_basic <- predict(cox_train_basic, newdata = test_data, type = "risk")

# Evaluate Concordance (C-index)
concordance(Surv(SurvivalMonths, Event) ~ risk_score_basic, data = test_data)

```

```{r}
# Split into training (70%) and testing (30%)
set.seed(380)
train_indices <- sample(seq_len(nrow(df)), size = 0.7 * nrow(df))
train_data <- df[train_indices, ]
test_data <- df[-train_indices, ]

# Refit the spline model on training data
cox_train <- coxph(Surv(SurvivalMonths, Event) ~ 
                     YearsSchooling + YearsWorkExperience + ExperienceSq +
                     IndustryExperience + SelfEmploymentExp + FounderFather +
                     Follower + Affiliated + 
                     ns(CapitalInvested, df = 3) + ns(PlanningDurationMonths, df = 3) +
                     EmployeesLog + RegisteredFirm + Specialist + Innovative + NationalMarket +
                     Munich + Branch_Construction + Branch_Transportation +
                     Branch_ComputerServices + Branch_OtherServices + WrittenBusinessPlan,
                   data = train_data)

# Risk score prediction
test_data$risk_score <- predict(cox_train, newdata = test_data, type = "risk")
concordance(Surv(SurvivalMonths, Event) ~ risk_score, data = test_data)
```

```{r}
# Visual table
knitr::kable(
  data.frame(
    Model = c("Cox Model (Basic)", "Cox Model (With Splines)"),
    Concordance_Index = c(0.2569, 0.2547),
    Std_Error = c(0.01903, 0.01897)
  ),
  caption = "Comparison of Predictive Performance Between Basic and Spline-Augmented Cox Models"
)
```

After splitting the dataset into training and testing subsets, we fitted two Cox proportional hazards models: a basic model assuming linear effects for all predictors, and an enhanced model incorporating restricted cubic splines (df = 3) for Capital Invested and Planning Duration Months. The concordance index (C-index) on the testing data for the basic model was 0.2569 (SE = 0.01903), while the spline-enhanced model achieved a similar C-index of 0.2547 (SE = 0.01897). Although the predictive accuracy, as measured by the C-index, did not significantly improve with splines, incorporating splines helped better capture potential non-linear relationships in the data, as suggested by earlier residual analyses. This adjustment improves model specification, reduces bias in coefficient estimates, and ensures the model more accurately reflects the true underlying relationships. Thus, even in the absence of major gains in predictive performance, the spline-augmented model is preferable for providing a more theoretically sound and robust analysis compared to the basic Cox model or earlier models in the original Brüderl et al. (1992) paper.

# Best Model Selection

```{r}
# Get predicted hazard values and deviance residuals
surv_model <- survreg(
  Surv(SurvivalMonths, Event) ~ 
    YearsSchooling + YearsWorkExperience + ExperienceSq +
    IndustryExperience + SelfEmploymentExp + FounderFather +
    Follower + Affiliated + CapitalLog + EmployeesLog + RegisteredFirm +
    Specialist + Innovative + NationalMarket + Munich +
    Branch_Construction + Branch_Transportation +
    Branch_ComputerServices + Branch_OtherServices +
    WrittenBusinessPlan + PlanningDurationMonths,
  data = df,
  dist = "loglogistic"
)

# Predicted linear predictor and deviance residuals
df$linear_predictor <- predict(surv_model, type = "lp")
df$deviance_resid <- residuals(surv_model, type = "deviance")

# Martingale residuals (using Cox model as approximation for these residuals)
cox_model <- coxph(
  Surv(SurvivalMonths, Event) ~ 
    YearsSchooling + YearsWorkExperience + ExperienceSq +
    IndustryExperience + SelfEmploymentExp + FounderFather +
    Follower + Affiliated + CapitalLog + EmployeesLog + RegisteredFirm +
    Specialist + Innovative + NationalMarket + Munich +
    Branch_Construction + Branch_Transportation +
    Branch_ComputerServices + Branch_OtherServices +
    WrittenBusinessPlan + PlanningDurationMonths,
  data = df
)

df$martingale_resid <- residuals(cox_model, type = "martingale")

# Plot martingale residuals against explanatory variables
vars_to_check <- c("YearsSchooling", "YearsWorkExperience", "ExperienceSq",
                   "IndustryExperience", "SelfEmploymentExp", "FounderFather",
                   "Follower", "Affiliated", "CapitalLog", "EmployeesLog",
                   "RegisteredFirm", "Specialist", "Innovative", "NationalMarket",
                   "Munich", "WrittenBusinessPlan", "PlanningDurationMonths")

plot_list <- lapply(vars_to_check, function(varname) {
  ggplot(df, aes(x = !!sym(varname), y = martingale_resid)) +
    geom_point(alpha = 0.3) +
    geom_smooth(method = "loess", color = "blue", se = FALSE) +
    labs(title = paste("Martingale Residuals vs", varname),
         x = varname, y = "Martingale Residuals") +
    theme_minimal()
})

do.call(grid.arrange, c(plot_list[1:6], ncol = 2))  # adjust slices as needed

```
## Selected Model: Log-Logistic Parametric Survival Model
We selected the log-logistic model fit via the flexsurvreg package as the final model for analyzing the survival of newly founded businesses. This decision was driven by both theoretical alignment with Brüderl et al. (1992) and empirical evidence from model diagnostics.

The log-logistic model is especially appropriate when the hazard rate is non-monotonic, which was evident in our life-table mortality curve — failure risk increased shortly after founding (likely due to resource depletion), peaked during adolescence, and then declined among surviving firms. This behavior reflects the “liability of adolescence” pattern discussed by Brüderl et al., and the shape parameter 𝑝>1 in our model confirms this curvature.

In contrast, Cox proportional hazards models assume a monotonically changing hazard (unless extended with time-varying covariates), and would therefore not capture this inverted U-shape without significant modifications. Moreover, our parametric model showed an excellent fit to the empirical hazard function, validating its suitability.

Residual Diagnostics
To assess the adequacy of model fit and identify any misspecification, we computed martingale residuals from a Cox model approximation (since flexsurvreg does not support these residuals directly). We plotted these residuals against all key covariates.

### Martingale Residual Plots

**Continuous Variables:**
YearsSchooling: Residuals show tight, symmetric dispersion around zero. The loess line is flat, suggesting a good linear fit. No transformation needed.

YearsWorkExperience & ExperienceSq: Residuals are well-behaved and evenly spread. Including both linear and squared terms was appropriate and captures the non-linear return to experience.

CapitalLog & EmployeesLog: Residuals are tightly clustered and show no systematic curvature — these log-transformations correctly normalized skewed data.

PlanningDurationMonths: Displays randomness around zero with no strong deviation, supporting linearity.

**Binary Variables:**
IndustryExperience, SelfEmploymentExp, FounderFather, Follower, Affiliated, RegisteredFirm, Specialist, Innovative, NationalMarket, Munich, WrittenBusinessPlan: These plots show the expected “stacked” appearance due to binary structure (residuals either high or low). The loess line is appropriately flat, indicating no strong interaction or omitted variable issues. No variables exhibit curvature or heteroscedasticity — an indication that functional form assumptions are appropriate.

The warnings above are not a concern for model diagnostics and do not affect residuals, model coefficients, or predictions. We retained loess lines for visual guidance but interpret them cautiously on binary plots.

## Strengths of the model

**Theoretical Alignment:** The log-logistic model closely aligns with established survival theory on firm dynamics, particularly the “liability of adolescence” described by Brüderl et al. (1992). It allows for non-constant hazard rates and flexible shapes, unlike more rigid exponential or Weibull models.

**Flexible Hazard Shape:** The estimated hazard function rises over time and begins to level off, capturing increasing but decelerating risk, which matches business lifecycle realities: firms are more vulnerable as they mature but may stabilize after initial challenges.

**Excellent Visual and Statistical Fit:** The predicted hazard and survival functions track closely with the observed patterns in your data. The survival curve’s 95% confidence band is tight, indicating a high level of precision. The significant shape parameter p>1 confirms a flexible, well-fitting model and allows interpretation of survival as an accelerating-decelerating process.

**Interpretability in AFT Framework:** The model's coefficients represent multiplicative effects on survival time, which are intuitive and useful for stakeholders (e.g., “each additional year of schooling multiplies survival time by X”).

**Diagnostics Support Model Validity:** Martingale residuals were centered, random, and pattern-free, indicating no major non-linearity or model misspecification. No transformations or interactions were missing based on visual diagnostics.

## Weaknesses

**Parametric Assumptions:** The model imposes a specific distributional form (log-logistic). While flexible, this still assumes a particular shape of survival times. If the true hazard is bimodal or varies erratically, this model may oversmooth or misrepresent the risk over time.

**Monotonic Hazard Limitation:** In your estimated hazard function, the curve increases and flattens — but does not show the classic inverted U-shape often associated with "liability of adolescence." While still reasonable, the model may be underestimating a possible early decline in risk for stabilized firms.

**Complexity for Lay Interpretation:** While coefficients in the AFT model are meaningful, interpreting effects on the hazard rate (vs. time) can be more abstract for non-technical stakeholders compared to simpler models like Kaplan-Meier curves.

**Residuals Based on Cox Approximation:** Because flexsurvreg doesn’t support martingale residuals, these were derived from a Cox model approximation. While acceptable, it's an indirect method, and diagnostic insights should be interpreted cautiously.

###Conclusion
Despite minor limitations, the log-logistic model demonstrates strong empirical fit, theoretical justification, and robust diagnostics, making it the best choice for analyzing the survival of newly founded businesses in this context. Its flexibility, statistical strength, and meaningful interpretation of time-based covariate effects make it a powerful tool for understanding early firm mortality.


# Executive Summary

### Problem Statement
Newly founded businesses face high uncertainty, particularly in their early stages. Understanding the factors that influence survival time is critical for improving the odds of long-term success. This analysis aims to identify and model the human capital, organizational, and environmental characteristics that significantly affect the duration of firm survival in Upper Bavaria, drawing on data inspired by Brüderl et al. (1992).

### Data Overview
The dataset used in this study originates from the Munich Founder Study and includes detailed information on 1,849 newly founded businesses registered in Upper Bavaria in 1985–1986. It contains 86 variables that capture a wide range of dimensions relevant to business survival.

*️ Variable Categorie**
The dataset contains the following major categories of variables:

Firm Demographics: StartYear, CompanyEndYear, SurvivalTime, Censored, EndYear_Clean

Industry and Environment: industry_category, Location, CompetitionIntensity, Seasonality, OrderClusters

Founder Characteristics (Human Capital): YearsSchooling, YearsWorkExperience, IndustryExperience, FounderFather, SelfEmploymentYears

Entrepreneurial Behavior (newly introduced variables): WrittenBusinessPlan (binary), PlanningDurationMonths (continuous)

Organizational Characteristics: LegalForm, CapitalInvested, EmployeeNum, FollowerBusiness, IndAffBusiness

Business Strategy: GenSpeBusiness (Generalist/Specialist), TradInnoBusiness (Traditional/Innovative), MarketScope (Local/National) Business

Time-Varying Variables (Unprocessed Episodes); beg1 to beg20, end1 to end20, typ1 to typ20: raw business episodes or transitions across time

Derived Variables for Modeling: SelfEmploymentExp (binary), ProblemType, and additional cleaned versions used in the final log-logistic model

**Preprocessing Highlights**
The original variables were cleaned and transformed to produce log-transformed capital and employee size, binary flags (e.g., registered firm, follower), and squared experience terms. A new binary outcome variable (Censored) was used to correctly define survival vs. right-censoring. Planning behavior (e.g., months spent preparing and business planning presence) was introduced as new explanatory variables, expanding on the original Brüderl framework.

### Modeling Approach
The log-logistic parametric survival model was selected as the best-fitting and most interpretable approach for this analysis. While Cox models with splines offer flexibility by allowing non-linear effects of time or covariates, they lack a fully specified functional form for the baseline hazard and can be more difficult to interpret—especially in small to medium samples. In contrast, the log-logistic model directly estimates a closed-form survival and hazard function, which is advantageous for interpretation and prediction. Moreover, the shape parameter in the log-logistic model confirms a non-monotonic, increasing hazard over time, aligning well with the “liability of adolescence” theory.

Although the Cox model with splines could approximate this shape, it introduces complexity without significantly improving fit, and does not yield intuitive time ratios for covariates as in the AFT framework. Diagnostic checks, including martingale residuals, survival curves, and the hazard function, confirmed that the log-logistic model provides a clean and accurate fit with no signs of misspecification. Thus, we prefer the parsimonious, theory-consistent, and easily interpretable structure of the log-logistic model over more flexible but opaque semi-parametric alternatives.

### Key Takeaways
This study successfully replicates several core findings from Brüderl et al. (1992): firms founded by individuals with higher levels of schooling, prior work experience, and industry-specific experience have significantly lower failure rates. Organizational characteristics such as greater initial capital, more employees, and a national market focus were also found to reduce failure risk, reaffirming the importance of startup resources and strategy.

In addition to validating these classic predictors, our model introduced and evaluated new variables not previously included in the Brüderl framework. Notably, the presence of a written business plan and a longer planning duration were associated with improved survival outcomes, highlighting the importance of founder preparedness and structured planning. These results suggest that beyond human and financial capital, strategic pre-launch behaviors play a critical role in business longevity — a valuable insight for modern entrepreneurial support programs and incubators.

**Replicated Findings:** Variables such as industry-specific experience, affiliation to other businesses, and having a father who was self-employed continued to show significant effects on survival, supporting Brüderl’s original findings.

**Human capital matters:** Years of schooling, work experience, and prior industry experience significantly increase survival time.

**Firm size and registration:** Larger initial capital, more employees, and being registered in the commercial register significantly reduce the risk of failure.

**Organizational strategy:** Businesses aiming at a national market outperform locally focused ones.

**Environment:** Retail and transportation firms have higher mortality rates, while manufacturing and construction show stronger survival.

### Future Directions
While the log-logistic model provided a strong fit for analyzing survival patterns among newly founded firms, several promising directions remain for future research that go well beyond refining methodology. First, it would be valuable to replicate this study using more recent datasets, especially to examine whether survival determinants observed in the 1980s still hold in today’s dramatically evolved entrepreneurial environment. The rise of digital business models, globalized startup ecosystems, and new financing structures (e.g., venture capital, crowdfunding) may have altered both risk exposure and success trajectories. Additionally, future studies could expand the outcome of interest beyond survival. By linking firm survival to profitability, growth, or job creation, we can differentiate between businesses that merely persist and those that thrive economically and contribute meaningfully to the economy.

On the modeling front, incorporating time-varying covariates (such as evolving competition, funding changes, or founder role shifts) would allow researchers to capture dynamic shifts in hazard risk over time. Likewise, integrating macroeconomic variables — such as interest rates, inflation, or financial crises — could better contextualize firm-level outcomes within broader economic cycles. There is also rich potential in deepening the behavioral and psychological lens. Variables like founder resilience, risk tolerance, and decision-making styles, captured via targeted surveys or even text mining of founder interviews, could enhance our understanding of the human capital drivers behind entrepreneurial survival.

From a methodological standpoint, machine learning techniques like random survival forests or DeepSurv neural networks could be explored to capture nonlinear interactions and high-dimensional relationships that traditional models may miss. While these approaches may trade off interpretability, they offer greater predictive power for practical applications like early-warning systems or funding decisions. Together, these future directions would not only enhance predictive performance but also enrich the theoretical understanding of entrepreneurship, providing a more holistic view of what it means for a firm not just to survive — but to succeed.

### Conclusion
This study applied a log-logistic parametric survival model to analyze the survival dynamics of newly founded firms in Upper Bavaria using data from the Munich Founder Study. The model was chosen over semi-parametric alternatives such as the Cox model with splines due to its ability to explicitly model a non-constant hazard, providing both a strong theoretical match to the concept of the "liability of adolescence" and a clean interpretability within the AFT framework. Model diagnostics, including martingale residuals, hazard and survival functions, and overall model fit, confirmed the log-logistic model’s suitability. Key findings validate and extend Brüderl et al.’s (1992) conclusions—highlighting the importance of human capital and startup scale—while also incorporating new predictors like planning duration and the presence of a business plan, which showed significant influence on firm longevity. These insights have direct implications for entrepreneurship training, incubator support programs, and public policy on small business development.